Parsing Postfix log files

### 2009-04-11 doc/logparser-conclusion.tex
Minor changes.
Corrections to the first coverage section.
Renamed Results chapter to Evaluation. Finished evaluation chapter introduction and summary.

@@

\newpage

>>>>

XXX INCLUDE A TIGHTLY FOCUSED FUTURE WORK SECTION\@.  PARSE LOG LINES FROM
OTHER PROGRAMS IF A WAY CAN BE FOUND TO ASSOCIATE THEM WITH EXISTING DATA
STRUCTURES\@.

@@

Experience implementing \parsername{} shows that full input coverage is
relatively easy to achieve with this architecture, and that with enough
time and effort a full understanding of the input is possible.  Postfix log
files would require substantial time and effort to correctly parse
regardless of the architecture used; this architecture enables an iterative
approach to be used (similar to Stepwise
Refinement~\cite{stepwise-refinement}), as is practiced in many other
software engineering disciplines.

### 2009-04-11 doc/logparser-implementation.tex
Renamed Results chapter to Evaluation. Finished evaluation chapter introduction and summary.

The framework collects data used when evaluating \parsernames{} efficiency
for the results chapter (\sectionref{Results}); some of the techniques used
to improve parsing speed can be turned off or altered to measure the effect
they have.  In each case the data gathered with the optimisation disabled
will be compared to the data gathered with it enabled, to quantify the
benefit each optimisation provides.  Three sets of data are collected:

@@

    \item The framework can skip inserting data into the database, because
        that dominates the execution time of the parser; the evaluation
        chapter measures the speed of \parsername{}, not the speed of the
        database or the disks it resides on.  The impact of storing results
        in the database is described at the beginning of
        \sectionref{Results}.

>>>>

The framework collects data used when evaluating \parsernames{} efficiency
for the evaluation chapter (\sectionref{Evaluation}); some of the
techniques used to improve parsing speed can be turned off or altered to
measure the effect they have.  In each case the data gathered with the
optimisation disabled will be compared to the data gathered with it
enabled, to quantify the benefit each optimisation provides.  Three sets of
data are collected:

@@

    \item The framework can skip inserting data into the database, because
        that dominates the execution time of the parser; the evaluation
        chapter measures the speed of \parsername{}, not the speed of the
        database or the disks it resides on.  The impact of storing results
        in the database is described at the beginning of
        \sectionref{Evaluation}.

### 2009-04-11 doc/logparser-introduction.tex
Renamed Results chapter to Evaluation. Finished evaluation chapter introduction and summary.

Chapter~\ref{Results} evaluates \parsernames{} efficiency, discussing the
various optimisations implemented in the framework, and the effect they
have.  It also discusses the coverage achieved by \parsername{} over
\numberOFlogFILES{} log files, with separate sections for the number of log
lines correctly recognised, and the number of connections and mails whose
journey through Postfix was correctly reconstructed.

@@

Portions of chapters~\ref{introduction} and~\ref{parser
architecture}--\ref{Results} have previously been published at a
conference~\cite{sgai-2008}, and later reprinted in a
journal~\cite{elsevier-2009}.  Publication of the conference paper was
supported by Science Foundation Ireland RFP~05/RF/CMS002.

>>>>

Chapter~\ref{Evaluation} evaluates \parsernames{} efficiency, discussing
the various optimisations implemented in the framework, and the effect they
have.  It also discusses the coverage achieved by \parsername{} over
\numberOFlogFILES{} log files, with separate sections for the number of log
lines correctly recognised, and the number of connections and mails whose
journey through Postfix was correctly reconstructed.

@@

Portions of chapters~\ref{introduction} and~\ref{parser
architecture}--\ref{Evaluation} have previously been published at a
conference~\cite{sgai-2008}, and later reprinted in a
journal~\cite{elsevier-2009}.  Publication of the conference paper was
supported by Science Foundation Ireland RFP~05/RF/CMS002.

### 2009-04-11 doc/logparser-results.tex
Renamed Results chapter to Evaluation. Finished evaluation chapter introduction and summary.

\chapter{Results And Evaluation}

\label{Results}

@@

XXX INTRODUCE COVERAGE\@.

@@

XXX IS THERE A BETTER WORD THAN ``coverage''?

XXX I THINK THIS SECTION NEEDS SUBSTANTIAL WORK, IF NOT A RE-WRITE\@; START
WITH THE PAPER'S CONTENT\@.

@@

    \item 9 mails whose initial log lines were not present in the log
        files.  XXX WHY WERE THERE NO WARNINGS ABOUT THESE\@?

@@

XXX EXTEND TO SUMMARISE PARSER EFFICIENCY TOO\@.

XXX REWRITE WHEN THIS CHAPTER IS FINISHED\@.

This section began with parser scalability, showing the linear relationship
between parsing time and input size.  Demonstrates the effect of rule
ordering on parsing time, and the unexpected consequences of specific
inputs.  The necessity of caching compiled regexes is attested to by the
third group of graphs, where the difference between caching and not caching
is staggering.  The penultimate section contains a breakdown of the rule
hits accumulated during a single test run of the parser.  Miscellaneous
graphs expected to be useful are collected in the final section.

XXX REWRITE COVERAGE SUMMARY IF NECESSARY\@.

Parser coverage is divided into two topics in this section: log lines
covered and mails covered.  The former is initially more important, as the
parser must successfully parse every line if it is to be complete, but
subsequently the latter takes precedence because reproducing the path a
mail takes through Postfix is the aim of the parser.  Increasing the
percentage of log lines parsed is relatively simple and non-intrusive:
adding new rules or modifying existing rules is simplified by the
separation of rules, actions, and framework.  Improving the logical
coverage is harder, as the actions taken by Postfix must be reconstructed
by the author, and the new sequence of actions integrated into the existing
model without breaking the existing parsing.  Detecting a deficiency in the
parsing algorithm is also significantly harder than detecting unparsed log
lines, as the parser will warn about any unparsed line, whereas discovering
a flaw in the parser requires understanding of the warnings produced and
the mails remaining in the state table.  Rectifying a flaw in the parser
requires an understanding of both the parser and Postfix's log files, and
investigative work to determine the cause of the deficiency, followed by
further examination of the log files in developing a solution.

Experience implementing the Postfix log parser shows that full input
coverage is relatively easy to achieve with this architecture, and that
with enough time and effort full understanding of the input is possible.
Postfix log files would require substantial time and effort to correctly
parse regardless of the architecture used; this architecture enables an
iterative approach to be used (similar to Stepwise
Refinement~\cite{stepwise-refinement}), as is practiced in many other
software engineering disciplines.

>>>>

\chapter{Evaluation}

\label{Evaluation}

@@

The second criterion the parser is evaluated on is its coverage of Postfix
log files.  This topic consists of two sections: what proportion of log
lines are correctly recognised by the ruleset, and what proportion of mail
delivery attempts are correctly understood and reconstructed by the
actions?  The former is a requirement before the latter can be achieved,
and the latter is important because the data provided by \parsername{} must
be both complete and correct for it to be of use to others.

@@

@@

    \item 9 mails whose initial log lines were not present in the log
        files.  There were no warnings for these mails because they were
        resemble child mails waiting to be tracked with a parent, see
        \sectionref{Re-injected mails} for details.

@@

This chapter evaluated \parsername{} on two criteria: efficiency, and
coverage of Postfix log files.  The former began by describing the mail
server the log files were taken from, the computer used to generate the
statistics in this chapter, and how the parser scales as the size of log
files increases, including why performance is better on the larger log
files.  The framework optimises the order in which rules are used when
trying to recognise each log line, and the effect that optimisation has is
explored; this is followed by the impact that adding more rules to the
ruleset has on parser performance.  The simplest and most effective of the
optimisations, caching compiled regexes, is described next, and the
efficiency evaluation concludes with an examination of where parsing time
is spent: recognising log lines or processing them?

Coverage of Postfix log files is divided into two topics in this chapter:
log lines correctly recognised, and mail delivery attempts correctly
understood and reconstructed.  The former is initially more important,
because the parser must correctly recognise every log line if it is to be
complete, but subsequently the latter takes precedence because correctly
reconstructing the journey a mail delivery attempt takes through Postfix is
the aim of the parser.  Increasing the proportion of log lines correctly
recognised is relatively simple and non-intrusive: adding new rules or
modifying existing rules is very easy because of the separation of rules,
actions, and framework.  Improving the understanding and reconstruction of
mail delivery attempts is harder, because Postfix's behaviour must be
analysed and figured out, and the new behaviour integrated into the actions
without breaking the existing parsing.  Detecting a deficiency in the
parser is also significantly harder, because the parser will warn about
unrecognised log lines, whereas discovering a flaw in the parser requires
careful study of any warnings produced and the entries remaining in the
state table.  Rectifying a flaw in the parser requires a deep understanding
of both the parser and Postfix's log files, investigative work to determine
the cause of the deficiency, further examination of the log files to aid in
developing a solution, and finally implementation, integration, and testing
of the solution.

This chapter shows that it is possible to balance the conflicting goals of
efficient and accurate parsing, and that one does not have to be sacrificed
to achieve the other.
